\section{Outline of paper}

Key research question: 

\emph{How do different sampling schemes affect system load and performance of WCA applications?}

Secondary question answered:

\emph{How can an artificial model of human behavior simplify reasearch in WCA?}

\begin{enumerate}
    \item Introduction
    \item Background and related work
        \begin{enumerate}
            \item What is WCA?
            \item Why is it interesting?
            \item Example applications and systems.
            \item Studying the impact of WCA on edge computing systems (why do we need a model for WCA?).
            \item Sampling strategy as an important factor driving system load.
            \item Impact of sampling schemes on system load.
            \item Include examples from other systems (control theory).
            \item Previous work on sampling in WCA.
            \item Research on sampling is hampered by the lack of a model for WCA.
            \item Previous approaches to understanding WCA behavior and modeling it.
        \end{enumerate}
    \item Model for WCA
        \begin{enumerate}
            \item Reiterate reasons for model
            \item Discuss base data from PLOS paper
                \begin{enumerate}
                    \item Timing
                    \item Frames
                \end{enumerate}
            \item Timing model
            \item Frame statistics
            \item Discuss effects of neuroticism (Bobby)
            \item Implementation in Python
        \end{enumerate}
    \item Validation of WCA Model
    \begin{enumerate}
        \item Reference model implementation (naive model)
        \item Plain comparison between naive, empirical, theoretical model using zero-wait sampling.
        \item Scaling behavior? 
    \end{enumerate}
    
    \item Sampling strategies
    \begin{enumerate}
        \item Reference strategies: zero-wait, ideal.
        \item Basic strategies: periodic, hold.
        \item Advanced strategy: adaptive aperiodic. Need to reference Vishnu's paper, indicate we're using a modified approximate solution.
    \end{enumerate}

    \item Experimental setup
    \begin{enumerate}
        \item Brief introduction to testbed
        \item Architectural design of experiment
    \end{enumerate}

    \item Results
    \item Discussion
    \item Conclusion
\end{enumerate}
\newpage